### 2015-05-26 paper.tex
paper -> revising 1

\documentclass{llncs}
\usepackage[utf8]{inputenc}
\usepackage[noend]{algorithmic}
\usepackage{algorithm}
\usepackage{epsfig}
\usepackage{multicol}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{subfigure}
\usepackage{url}
\newcommand{\verbatimproperties}{\renewcommand{\baselinestretch}{0.85} \small}
%\newcommand{\tableproperties}{\centering \big}
\newcommand{\LINECOMMENT}[1]{$\triangleright$ #1}
%\setlength{\belowcaptionskip}{-26pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\title{On Extending a Full-Sharing Multithreaded Tabling Design with
  Batched Scheduling}

\author{Miguel Areias \and Ricardo Rocha}

\institute{CRACS \& INESC TEC and Faculty of Sciences, University of Porto\\
           Rua do Campo Alegre, 1021, 4169-007 Porto, Portugal\\
           \email{\{miguel-areias,ricroc\}@dcc.fc.up.pt}}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstract}
Tabling is an implementation technique that overcomes some limitations
of traditional Prolog systems in dealing with redundant
sub-computations and recursion. Multithreading in Prolog is the
ability to concurrently perform computations, in which each
computation runs independently but shares the program clauses. When
multithreading is combined with tabling, we have the best of both
worlds, since we can exploit the combination of higher procedural
control with higher declarative semantics. To support this
combination, the Yap Prolog system has, at engine level, multiple
designs that vary from a No-Sharing design, where each thread
allocates fully private tables for each new tabled subgoal called
during its computation, to a Full-Sharing design, where threads share
the complete table space. Arguably, batched scheduling is one of the
most successful tabling scheduling strategies. In this work, we
propose the PAC strategy to combine the Full-Sharing design with
batched scheduling. Experimental results show that, PAC is a good
first approach, since with it the FS design remains quite
competitive.\\

\textbf{Keywords:} Logic Programming, Multithreading, Tabling, Scheduling.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}

Tabling~\cite{Chen-96} is an implementation technique that overcomes
some limitations of traditional Prolog systems in dealing with
redundant sub-computations and recursion. Tabling consists of storing
intermediate answers for subgoals in a proper data structure, called
the \emph{table space}, so that they can be reused when a repeated
subgoal appears during the resolution process. Tabling has become a
popular and successful technique thanks to the ground-breaking work in
the XSB Prolog system and in particular in the SLG-WAM
engine~\cite{Sagonas-98}, the most successful engine of XSB. The
success of SLG-WAM led to several alternative implementations that
differ in the execution rule, in the data-structures used to implement
tabling, and in the changes to the underlying Prolog
engine. Implementations of tabling are now widely available in systems
like Yap Prolog, B-Prolog, ALS-Prolog, Mercury, Ciao Prolog and more
recently Picat.

Multithreading in Prolog is the ability to concurrently perform
computations, in which each computation runs independently but shares
the program clauses. When multithreading is combined with tabling, we
have the best of both worlds, since we can exploit the combination of
higher procedural control with higher declarative semantics. To the
best of our knowledge, XSB~\cite{Marques-08} and Yap~\cite{Areias-12a}
are the only Prolog systems that support the combination of
multithreading with tabling. Yap implements a SWI-Prolog compatible
multithreading library~\cite{Wielemaker-03}. Yap's threads have their
own execution stacks and only share the code area where predicates,
records, flags and other global non-backtrackable data are stored. For
tabled evaluation, a thread views its tables as private but, at the
engine level, Yap has three designs~\cite{Areias-12a} that vary from a
\emph{No-Sharing} (NS) design, where each thread allocates fully
private tables for each new tabled subgoal called during its
computation, to a \emph{Full-Sharing} (FS) design, where threads share
the complete table space.

The decision about the evaluation flow is determined by the
\emph{scheduling strategy}. Different strategies may have a
significant impact on performance, and may lead to a different
ordering of solutions to the query goal. Arguably, the two most
successful tabling scheduling strategies are \emph{local scheduling}
and \emph{batched scheduling}~\cite{Freire-96}. Local scheduling tries
to complete subgoals as soon as possible. When new answers are found,
they are added to the table space and the evaluation fails. Answers
are only returned when all program clauses for the subgoal at hand
were resolved, i.e. after the complete evaluation of subgoal's search
space. Local scheduling has the advantage of minimizing the size of
clusters of dependent subgoals, however it delays propagation of
answers and requires the complete evaluation of the search
space.

Batched scheduling favors forward execution first, backtracking next,
and consuming answers or completion last. It thus tries to delay the
need to move around the search tree by batching the return of answers
to repeated subgoals. When new answers are found for a particular
tabled subgoal, they are added to the table space and the evaluation
continues. Batched scheduling can be an useful strategy in tabled
logic programs that require an eager propagation of answers and/or do
not require the complete set of answers to be found.

With the FS design, all tables are shared. Thus, since several threads
can be inserting answers in the same table, when an answer already
exists, it is not possible to determine if the answer is new or
repeated for a certain thread without further support. However, for
local scheduling this is not a problem because, for repeated and new
answers, local scheduling always fails. The problem is with batched
scheduling that requires further support to work with the FS design
since, with batched scheduling, only repeated answers should fail. We
have thus to detect when a newly found answer is also a new answer for
the thread at hand in order to ensure that new answers for a given
subgoal call are correctly propagated one and only once to all
thread's repeated calls.

In this work, we propose an extension to the table space data
structures, which we named \emph{Private Answer Chaining (PAC)}, as a
way to keep track, per thread and subgoal call, of the answers that
were already found and propagated to all repeated calls. We discuss in
detail our proposal for extending the FS design with batched
scheduling and we present a performance analysis comparison between
local and batched scheduling. Experimental results show that, despite
the extra PAC data structures required to support batched scheduling
with the FS design, the execution time of the combination is still
quite competitive when comparing against the FS design with local
scheduling.

The remainder of the paper is organized as follows. First, we briefly
introduce some background and related work. Then, we describe our
approach for combining batched scheduling with the FS design and we
discuss the most important implementation details. Finally, we present
experimental results and we end by outlining some conclusions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Background}

This section introduces some background needed for the following
sections.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Table Space Organization}

The basic idea behind tabling is straightforward: programs are
evaluated by storing answers for tabled subgoals in an appropriate
data space, called the \emph{table space}. Similar calls to tabled
subgoals are not re-evaluated against the program clauses, instead
they are resolved by consuming the answers already stored in their
table entries. During this process, as further new answers are found,
they are stored in their tables and later returned to all similar
calls.

\begin{wrapfigure}{r}{4.5cm}
\vspace{-\intextsep}
\centering
\includegraphics[width=4.5cm]{figures/table-space.pdf}
\caption{Table space organization}
\label{fig_table_space}
\vspace{-0.5\intextsep}
\end{wrapfigure}

Figure~\ref{fig_table_space} shows the table space organization of the
Yap system. At the entry point we have the \emph{table entry} data
structure. This structure is allocated when a tabled predicate is
being compiled, so that a pointer to the table entry can be included
in its compiled code. This guarantees that further calls to the
predicate will access the table space starting from the same
point. Below the table entry, we have the \emph{subgoal trie
  structure}. Each different tabled subgoal call to the predicate at
hand corresponds to a unique path through the subgoal trie structure,
always starting from the table entry, passing by several subgoal trie
data units, the \emph{subgoal trie nodes}, and reaching a leaf data
structure, the \emph{subgoal frame}. The subgoal frame stores
additional information about the subgoal and acts like an entry point
to the \emph{answer trie structure}. Each unique path through the
answer trie data units, the \emph{answer trie nodes}, corresponds to a
different answer to the entry subgoal.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Yap's Multithreaded Tabling Support}

In Yap, a thread views its tables as private but, at the engine level,
it implements three designs for concurrent tabling support that vary
from a \emph{No-Sharing} (NS) design, where each thread allocates
fully private tables for each new tabled subgoal called during its
computation, to a \emph{Full-Sharing} (FS) design, where threads share
the complete table space. Figure~\ref{fig_yap_mt_support} shows Yap's
multithreaded table space organization for the NS and FS designs,
where an interface layer abstracts the design being used at the engine
level. The figure illustrates the main differences between the two
designs for a situation where several threads are evaluating the same
tabled subgoal call $call\_i$.

\begin{figure}[!ht]
%\vspace{-0.5\intextsep}
\centering
\includegraphics[width=9.5cm]{figures/yap-mt.pdf}
\caption{Yap's multithreaded table space organization for the NS and FS designs}
\label{fig_yap_mt_support}
%\vspace{-1.0\intextsep}
\end{figure}

When using the NS design, each thread allocates fully private tables
for each new subgoal called during its computation. One can observe
that the table entry data structure still stores the common
information for the predicate (such as the arity or the evaluation
strategy), and then each thread $t$ has its own cell $T_t$ inside a
\emph{bucket array} which points to the private data structures.

When using the FS design, the subgoal and answer trie structures and
part of the subgoal frame information (the \emph{subgoal entry} data
structure in Fig.~\ref{fig_yap_mt_support}) are shared among all
threads. The previous subgoal frame data structure was split in two:
the \emph{subgoal entry} stores common information for the subgoal
call (such as the pointer to the shared answer trie structure); the
remaining information is kept private to each thread in the
\emph{subgoal frame} data structure.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Scheduling Strategies}

Local scheduling schedules the evaluation of a program in a
breath-first manner. It favors backtracking first with completion
instead of forward execution, leaving the consumption of answers for
last. Local scheduling only allows a \emph{Cluster of Dependent
  Subgoals} (CDS) to return answers after a fix-point has been
reached~\cite{Freire-96}. In other words, local scheduling tries to
keep a CDS as minimal as possible, thus creating less complex
dependencies between subgoals, which causes a sooner completion of
subgoals.

On the other hand, batched scheduling schedules the evaluation of a
program in a depth-first manner. It favors the forward execution first
instead of backtracking, leaving the consumption of answers and
completion for last. It thus tries to delay the need to move around
the search tree by batching the return of answers. When new answers
are found for a particular tabled subgoal, they are added to the table
space and the execution continues. For some situations, this results
in creating dependencies to older subgoals, therefore enlarging the
current CDS and delaying the fix-point that guarantees that all
dependent subgoals in a CDS are completely
evaluated~\cite{Sagonas-98}. Batched scheduling can be an useful
strategy in tabled logic programs that require an eager propagation of
answers and/or do not require the complete set of answers to be found
(partial evaluation of the search space).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Extending Full-Sharing with Batched Scheduling}

In this section, we describe our proposal to support the combination
of batched scheduling with the FS design. Remember that batched
scheduling requires further support to work with the FS design
because, with batched scheduling, we have to detect when a newly found
answer is also a new answer for the thread at hand in order to ensure
that new answers for a given subgoal call are correctly propagated one
and only once to all thread’s repeated calls. In the FS design answer
representation and answer propagation is stored in the answer trie
data structure, thus threads are unable to know if they have or not
have propagated an answer whenever it appears during the
evaluation. To solve that, we propose an extension to the table space
data structures, which we named \emph{Private Answer Chaining} (PAC),
as a way to keep track, per thread and subgoal call, of the answers
that were already found and propagated to the thread's repeated
calls. Figure~\ref{fig_tabtries_pcc_key-idea} illustrates the key idea
of the PAC strategy. In a nutshell, PAC splits answer representation
from answer propagation, and allows the first to be kept publicly
shared among threads in the answer trie data structure, and the second
to be privately stored in the subgoal frame data structure of each
thread. Thus, with PAC, whenever a answer appears during the
evaluation, a thread knows if the answer was already propagated or not
propagated, just by consulting its own subgoal frame data structure.

\begin{wrapfigure}{r}{6.5cm}
\vspace{-\intextsep}
\centering
\includegraphics[width=6.5cm]{figures/pcc_key-idea.pdf}
\caption{PAC splits answer representation and answer propagation}
\label{fig_tabtries_pcc_key-idea}
\vspace{-0.5\intextsep}
\end{wrapfigure}

%% \begin{figure}[!ht]
%% \centering
%% \includegraphics[width=7.5cm]{figures/pcc_key-idea.pdf}
%% \caption{PAC splits answer representation and answer propagation}
%% \label{fig_tabtries_pcc_key-idea}
%% \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Our Approach}

The PAC procedure works at the subgoal frame level, which is private
to each thread. The key idea of the PAC data structure is then to
extend subgoal frames with an auxiliary private chaining of answers
for each subgoal call, in order to keep track of the answers already
found for the call. Later, if a thread completes a subgoal's
evaluation, i.e, if the subgoal's table is marked as complete, its PAC
is made public, so that from that point on all threads can use that
chain in complete (only reading) mode. Figure~\ref{fig_tabtries_pcc}
illustrates the new data structures involved in the implementation of
our PAC's proposal for a situation where several threads are
evaluating the same tabled subgoal call $call\_i$.

\begin{figure}[!ht]
%\vspace{-\intextsep}
\centering
\includegraphics[width=10.5cm]{figures/pcc.pdf}
\caption{PAC's data structures for (a) private and (b) public
  chaining}
\label{fig_tabtries_pcc}
%\vspace{-1.0\intextsep}
\end{figure}

Figure~\ref{fig_tabtries_pcc}(a) shows then a situation where two
threads, $T_1$ and $T_{k-1}$, are sharing the same subgoal entry for
the call $call\_i$ when the subgoal is still under evaluation, i.e.,
the subgoal is not yet complete. The current state of the evaluation
shows an answer trie with 3 answers found for the subgoal call
$call\_i$. For the sake of simplicity, we are omitting the internal
answer trie nodes and we are only showing the leaf nodes nodes $LN_1$,
$LN_2$ and $LN_3$, in the figure.

With PAC support, the leaf nodes are not chained in the answer trie
data structure, as usual. Now, the chaining process is done privately,
and for that, we use the subgoal frame structure of each thread. On
the subgoal frame structure we added a new field, called
\emph{Answers}, to store the answers found within the execution of the
thread. In order to minimize PAC's impact, each answer node in the
private chaining has only two fields: (i) an entry pointer, which
points to the corresponding leaf node in the answer trie data
structure; and (ii) a next pointer to chain the nodes in the private
chaining. To maintain a good performance, when the number of answer
nodes exceeds a certain threshold, we use a hash trie mechanism design
similar to the one presented in the
work~\cite{Areias-ijpp15}. However, since this mechanism is private to
each thread, it does not require any concurrency support.

Going back to Fig.~\ref{fig_tabtries_pcc}(a), the PAC's data
structures represent then two different situations where threads can
be evaluating a subgoal call. Thread $T_1$ has only found one answer
and it is using a direct answer chaining to access the node
$LN_1$. Thread $T_{k-1}$ was already found three answers for the
subgoal call and it is using the hash trie mechanism within its
private chaining. Please note that the answer nodes are still chained
between themselves, thus that repeated calls belonging to thread
$T_{k-1}$ can consume the answers as in the original mechanism.

Figure~\ref{fig_tabtries_pcc}(b) shows the state of the subgoal call
after completion. When a thread $T$ completes a subgoal call, it frees
its private consumer structures, but before doing that, it checks
whether another thread as already marked the subgoal as completed. If
no other thread has done that, then thread $T$ not only follows its
private chaining mechanism as it would for freeing its private nodes,
but also, follows the pointers to the answer trie leaf nodes in order
to create a chain inside the answer trie. Since this procedure is done
inside a critical region, no more than one thread can be doing this
chaining process. Thus, in Fig.~\ref{fig_tabtries_pcc}(b), we are
showing the situation where the subgoal call is completed and both
threads $T_1$ and $T_{k-1}$ have already chained the leaf nodes inside
the answer trie and removed their private chaining structures.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Implementations Details}

At the tabling engine, the major difference between local and batched
scheduling is in the tabling operation \emph{tabled new answer}, where
we decide what to do when an answer is found during the
evaluation. This operation checks whether a newly found answer is
already in the corresponding answer trie structure and, if not,
inserts it. Algorithm~\ref{alg_table_new_answer_batched} shows how we
have extended this operation to support the FS design with batched
scheduling.

\begin{algorithm} [!ht]
\caption{tabled\_new\_answer(answer ANS, subgoal frame SF)}
\algsetup{indent=0.65cm}
\begin{algorithmic}[1]
  \STATE $leaf \gets check\_insert\_answer\_trie(ANS, SF)$
  \STATE $chain \gets check\_insert\_consumer\_chain(leaf, SF)$
  \IF {$is\_answer\_marked\_as\_found(chain) = True$}
    \RETURN $failure$
  \ELSE                                        [the answer is new]
    \STATE $mark\_answer\_as\_found(chain)$
    \IF {$local\_scheduling\_mode(SF)$}
      \RETURN $failure$
    \ELSE                                [batched scheduling mode]
      \RETURN $proceed$
    \ENDIF
  \ENDIF  
\end{algorithmic}
\label{alg_table_new_answer_batched}
\end{algorithm}

The algorithm receives two arguments: the newly answer found during
the evaluation (\emph{ANS}) and the subgoal frame which corresponds to
the call at hand (\emph{SF}). The algorithm begins by
checking/inserting the given \emph{ANS} into the answer trie
structure, which will return the leaf node for the path representing
\emph{ANS} (line 1). Then, it checks/inserts the given \emph{leaf}
node into the private chaining for the current thread, which will
return the corresponding answer chain node (line 2). Next in line 3,
it tests whether the answer chain node already existed in the chain,
i.e., if it was inserted or not by the current check/insert operation
in order to return failure (line 4), or it proceeds with marking the
answer \emph{ANS} has found (line 6). At the end (lines 7 to 10), it
returns failure if local scheduling is active (line 8), otherwise,
batched scheduling is active, and it proceeds by propagating the
answer \emph{ANS} to the current execution environment (line 10).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Performance Analysis}

We now present experimental results about the usage of PAC in the FS
design with batched scheduling. Since without PAC the FS design would
not be able to be used with batched scheduling, to put results in
perspective, we will be showing also the results for local scheduling
and for the NS design. The environment for our experiments was a
machine with 32-Core AMD Opteron (TM) Processor 6274 (2 sockets with
16 cores each) with 32G of main memory, running the Linux kernel is
the 3.16.7-200.fc20.x86\_64 with Yap Prolog 6.3. For the experiments,
we used the \emph{TabMalloc} memory allocator~\cite{Areias-12b} and
the set of benchmarks described in~\cite{Areias-12b}. These benchmarks
create \emph{worst case scenarios}, where we are able to show the
lowest bounds of performance that each design might achieve when
applied/used in other real world applications/programs.

Table~\ref{tab_batched_overhead} shows the overhead ratios, when
compared with the NS design with 1 thread (running with local
scheduling without TabMalloc), for the NS and FS designs, when running
1, 8, 16, 24 and 32 threads, with TabMalloc and, local scheduling
(column \emph{Local}) and batched scheduling (column \emph{Batched})
strategies. In order to give a fair weight to each benchmark, the
overhead ratio is calculated as follows. We begin by running ten times
each benchmark $B$ for each design $D$ with $T$ threads. Then, we
calculate the average of those ten runs and use that value ($D_{BT}$)
to put it in perspective against the base time, which is the average
of the ten runs of the NS design with $1$ thread ($NS_{B1}$). For
that, we use the following formula for the overhead $O_{DBT} = D_{BT}
/ NS_{B1}$. After calculating all the overheads $O_{DBT}$ for a
certain design $D$ and number of threads $T$ corresponding to the
several benchmarks $B$, we calculate the respective minimum, average,
maximum and standard deviation overhead ratios.

\setlength{\tabcolsep}{12pt}
\begin{table}[!ht]
\centering
\caption{Overhead ratios, when compared with the NS design with $1$
  thread (running with local scheduling and without TabMalloc) for the NS
  and FS designs (with TabMalloc), when running 1, 8, 16, 24 and 32
  threads with local and batched scheduling (best ratios by row and by
  design for the Minimum, Average and Maximum are in bold)}
\begin{tabular}{ll||cc|cc}
\multicolumn{2}{c||}{\multirow{2}{*}{\bf Threads}} &
\multicolumn{2}{c|}{\multirow{1}{*}{\bf NS}} &
\multicolumn{2}{|c}{\multirow{1}{*}{\bf FS}}\\ \cline{3-6}
& 
& \multicolumn{1}{c}{\bf Local}
& \multicolumn{1}{c}{\bf Batched}
& \multicolumn{1}{|c}{\bf Local}
& \multicolumn{1}{c}{\bf Batched (PAC)}\\
\hline\hline
\multirow{4}{*}{\bf 1}
& {\bf Min }& {\bf 0.53}& 0.55& 1.01& {\bf 0.95}\\
& {\bf Avg }& {\bf 0.78}& 0.82& {\bf 1.30}& 1.46\\
& {\bf Max }& 1.06& {\bf 1.05}& {\bf 1.76}& 2.33\\
& {\bf StD }& 0.15& 0.14& 0.22& 0.44\\
\hline
\multirow{4}{*}{\bf 8}
& {\bf Min }& 0.66& {\bf 0.63}& 1.16&{\bf  0.99}\\
& {\bf Avg }& {\bf 0.85}& 0.88& {\bf 1.88}& 1.95\\
& {\bf Max }& {\bf 1.12}& 1.14& {\bf 2.82}& 3.49\\
& {\bf StD }& 0.13& 0.14& 0.60& 0.79\\
\hline
\multirow{4}{*}{\bf 16}
& {\bf Min }& 0.85& {\bf 0.75}& 1.17& {\bf 1.06}\\
& {\bf Avg }& {\bf 0.98}& 1.00& {\bf 1.97}& 2.08\\
& {\bf Max }& {\bf 1.16}& 1.31& {\bf 3.14}& 3.69\\
& {\bf StD }& 0.09& 0.17& 0.65& 0.83\\
\hline
\multirow{4}{*}{\bf 24}
& {\bf Min }& {\bf 0.91}& 0.93& 1.16& {\bf 1.09}\\
& {\bf Avg }& {\bf 1.15}& 1.16& {\bf 2.06}& 2.19\\
& {\bf Max }& 1.72& {\bf 1.60}& {\bf 3.49}& 4.08\\
& {\bf StD }& 0.20& 0.21& 0.70& 0.91\\
\hline
\multirow{4}{*}{\bf 32}
& {\bf Min }& 1.05& {\bf 1.04}& 1.33& {\bf 1.26}\\
& {\bf Avg }& 1.51& {\bf 1.49}& {\bf 2.24}& 2.41\\
& {\bf Max }& {\bf 2.52}& 2.63& {\bf 3.71}& 4.51\\
& {\bf StD }& 0.45& 0.45& 0.74& 1.02\\
\end{tabular}
\label{tab_batched_overhead}
\end{table}

By observing Table~\ref{tab_batched_overhead}, we can see that, for 1
thread, on the minimum overhead ratio, local scheduling is better on
the NS design, while batched scheduling is better in the FS
design. For the average and maximum overhead ratios, in the FS design,
the best strategy is always the local scheduling. Additionally, the
difference between local and batched scheduling in the FS is higher
than in the NS design, which indicates the overhead that the PAC
strategy brings to the FS design. Recall, whenever an answer is found
during the evaluation, PAC requires threads to traverse their private
consumer data structures to check if the answer was already
propagated. This is the reason for difference in the overheads that we
are observing in the experimental results. As we scale the number of
threads, for the NS design both scheduling strategies have similar
minimum, average and maximum overhead ratios. For the FS design, the
best minimum overhead ratio is always for the batched scheduling. For
the average and maximum overhead ratio, local scheduling is always
better than batched scheduling. However, for the average ratio, the
results between both strategies is quite close. For the maximum
overhead ratio, the difference between both strategies is
significantly higher than in the NS design. In conclusion,
experimental results show that on average, the PAC strategy does not
seem to have a big impact in the performance, however it still leaves
room for further improvements. The difference between both local and
batched scheduling is significantly higher in the FS design than in
the NS design.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusions and Further Work}

Local and batched scheduling are arguably two of the most well-known
tabling scheduling techniques. The major difference between them is
that local propagates answers only after all answer are found, while
batched propagates answers immediately after they are found. Batched
scheduling is a useful strategy in tabled logic programs that require
an eager propagation of answers and/or do not require the complete set
of answers to be found (partial evaluation of the search space). In
this work, we have presented the PAC strategy, which is a simple and
novel approach for combining the FS design with batched
scheduling. PAC splits answer representation from answer propagation,
and allows the first to be publicly shared among threads while the
second to be private to each thread.

Experimental results in worst-case scenarios showed that on average,
the PAC strategy does not seem to have a big impact in the
performance, however it still leaves room for further improvements
specially in the extra structures required to control the propagated
answers. Further work will include the usage of time-stamped tries to
minimize the search for the propagated answers and new real-world
problems that will allow us to improve and consolidate our framework.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Acknowledgments}

This work is funded by the project SIBILA (NORTE-07-0124-FEDER-000059)
under the North Portugal Regional Operational Programme (ON.2 – O Novo
Norte) and under the National Strategic Reference Framework (NSRF),
through the European Regional Development Fund (ERDF), and by national
funds, through the Portuguese funding agency, Fundação para a Ciência
e a Tecnologia (FCT).

%% This work is partially funded by the ERDF (European Regional
%% Development Fund) through the COMPETE Programme and by FCT (Portuguese
%% Foundation for Science and Technology) within projects SIBILA
%% (NORTE-07-0124-FEDER-000059) and PEst
%% (FCOMP-01-0124-FEDER-037281). Miguel Areias is funded by the FCT grant
%% SFRH/BD/69673/2010.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{splncs}
\bibliography{references}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}